---
title: "TV_Mini_Data-Analysis_Deliverable_1"
output: github_document
---

# Welcome to your (maybe) first-ever data analysis project!

And hopefully the first of many. Let's get started:

1.  Install the [`datateachr`](https://github.com/UBC-MDS/datateachr) package by typing the following into your **R terminal**:

<!-- -->

    install.packages("devtools")
    devtools::install_github("UBC-MDS/datateachr")

2.  Load the packages below.

```{r}
library(datateachr)
library(tidyverse)
```

3.  Make a repository in the <https://github.com/stat545ubc-2022> Organization. You will be working with this repository for the entire data analysis project. You can either make it public, or make it private and add the TA's and Lucy as collaborators. A link to help you create a private repository is available on the #collaborative-project Slack channel. 

# Instructions

## For Both Milestones

-   Each milestone is worth 45 points. The number of points allocated to each task will be annotated within each deliverable. Tasks that are more challenging will often be allocated more points.

-   10 points will be allocated to the reproducibility, cleanliness, and coherence of the overall analysis. While the two milestones will be submitted as independent deliverables, the analysis itself is a continuum - think of it as two chapters to a story. Each chapter, or in this case, portion of your analysis, should be easily followed through by someone unfamiliar with the content. [Here](https://swcarpentry.github.io/r-novice-inflammation/06-best-practices-R/) is a good resource for what constitutes "good code". Learning good coding practices early in your career will save you hassle later on!

## For Milestone 1

**To complete this milestone**, edit [this very `.Rmd` file](https://raw.githubusercontent.com/UBC-STAT/stat545.stat.ubc.ca/master/content/mini-project/mini-project-1.Rmd) directly. Fill in the sections that are tagged with `<!--- start your work below --->`.

**To submit this milestone**, make sure to knit this `.Rmd` file to an `.md` file by changing the YAML output settings from `output: html_document` to `output: github_document`. Commit and push all of your work to the mini-analysis GitHub repository you made earlier, and tag a release on GitHub. Then, submit a link to your tagged release on canvas.

**Points**: This milestone is worth 45 points: 43 for your analysis, 1 point for having your Milestone 1 document knit error-free, and 1 point for tagging your release on Github.

# Learning Objectives

By the end of this milestone, you should:

-   Become familiar with your dataset of choosing
-   Select 4 questions that you would like to answer with your data
-   Generate a reproducible and clear report using R Markdown
-   Become familiar with manipulating and summarizing your data in tibbles using `dplyr`, with a research question in mind.

# Task 1: Choose your favorite dataset (10 points)

The `datateachr` package by Hayley Boyce and Jordan Bourak currently composed of 7 semi-tidy datasets for educational purposes. Here is a brief description of each dataset:

-   *apt_buildings*: Acquired courtesy of The City of Toronto's Open Data Portal. It currently has 3455 rows and 37 columns.

-   *building_permits*: Acquired courtesy of The City of Vancouver's Open Data Portal. It currently has 20680 rows and 14 columns.

-   *cancer_sample*: Acquired courtesy of UCI Machine Learning Repository. It currently has 569 rows and 32 columns.

-   *flow_sample*: Acquired courtesy of The Government of Canada's Historical Hydrometric Database. It currently has 218 rows and 7 columns.

-   *parking_meters*: Acquired courtesy of The City of Vancouver's Open Data Portal. It currently has 10032 rows and 22 columns.

-   *steam_games*: Acquired courtesy of Kaggle. It currently has 40833 rows and 21 columns.

-   *vancouver_trees*: Acquired courtesy of The City of Vancouver's Open Data Portal. It currently has 146611 rows and 20 columns.

**Things to keep in mind**

-   We hope that this project will serve as practice for carrying our your own *independent* data analysis. Remember to comment your code, be explicit about what you are doing, and write notes in this markdown document when you feel that context is required. As you advance in the project, prompts and hints to do this will be diminished - it'll be up to you!

-   Before choosing a dataset, you should always keep in mind **your goal**, or in other ways, *what you wish to achieve with this data*. This mini data-analysis project focuses on *data wrangling*, *tidying*, and *visualization*. In short, it's a way for you to get your feet wet with exploring data on your own.

And that is exactly the first thing that you will do!

1.1 Out of the 7 datasets available in the `datateachr` package, choose **4** that appeal to you based on their description. Write your choices below:

**Note**: We encourage you to use the ones in the `datateachr` package, but if you have a dataset that you'd really like to use, you can include it here. But, please check with a member of the teaching team to see whether the dataset is of appropriate complexity. Also, include a **brief** description of the dataset here to help the teaching team understand your data.

<!-------------------------- Start your work below ---------------------------->

1: *steam_games*
2: *apt_buildings*
3: *building_permits*
4: *parking_meters*

<!----------------------------------------------------------------------------->

1.2 One way to narrowing down your selection is to *explore* the datasets. Use your knowledge of dplyr to find out at least *3* attributes about each of these datasets (an attribute is something such as number of rows, variables, class type...). The goal here is to have an idea of *what the data looks like*.

*Hint:* This is one of those times when you should think about the cleanliness of your analysis. I added a single code chunk for you below, but do you want to use more than one? Would you like to write more comments outside of the code chunk?

<!-------------------------- Start your work below ---------------------------->

```{r}
### EXPLORE HERE ###
#steam_games
class(steam_games)
head(steam_games)
summary(steam_games)
glimpse(steam_games)

#apt_buildings
class(apt_buildings)
head(apt_buildings)
summary(apt_buildings)
glimpse(apt_buildings)

#building_permits
class(building_permits)
head(building_permits)
summary(building_permits)
glimpse(building_permits)

#parking_meters
class(parking_meters)
head(parking_meters)
summary(parking_meters)
glimpse(parking_meters)

```

<!----------------------------------------------------------------------------->

1.3 Now that you've explored the 4 datasets that you were initially most interested in, let's narrow it down to 2. What lead you to choose these 2? Briefly explain your choices below, and feel free to include any code in your explanation.

<!-------------------------- Start your work below ---------------------------->

Choice 1: *steam_games*
Reason: 
- I am familiar with this data set, and understand what each variables and observations stand for.

Choice 2: *apt_buildings*
Reason: 
- Compare to other data set, this data set does not seem to have many missing or NA values
```{r}
apt_buildings %>% 
  select(id:no_barrier_free_accessible_units) %>%
  is.na() %>% 
  sum()
```
- Variables in this data set seems straightforward compared to the first data set
- I am interested in the information contained in this data set

<!----------------------------------------------------------------------------->

1.4 Time for the final decision! Going back to the beginning, it's important to have an *end goal* in mind. For example, if I had chosen the `titanic` dataset for my project, I might've wanted to explore the relationship between survival and other variables. Try to think of 1 research question that you would want to answer with each dataset. Note them down below, and make your final choice based on what seems more interesting to you!

<!-------------------------- Start your work below ---------------------------->

-   **Choice 1:** *steam_games*
*Research Question:* What is the relationship between "genre" and the average rating of review the receuved? 

-   **Choice 2:** *apt_buildings*
*Research Question*: What is the relationship between "air_conditioning " and "year_built"?  
--> This would be my final choice.I am interested in knowing that as the average summer in the last few years have been getting hotter, do more recent apartment equipped with air conditioning.

<!----------------------------------------------------------------------------->

# Important note

Read Tasks 2 and 3 *fully* before starting to complete either of them. Probably also a good point to grab a coffee to get ready for the fun part!

This project is semi-guided, but meant to be *independent*. For this reason, you will complete tasks 2 and 3 below (under the **START HERE** mark) as if you were writing your own exploratory data analysis report, and this guidance never existed! Feel free to add a brief introduction section to your project, format the document with markdown syntax as you deem appropriate, and structure the analysis as you deem appropriate. Remember, marks will be awarded for completion of the 4 tasks, but 10 points of the whole project are allocated to a reproducible and clean analysis. If you feel lost, you can find a sample data analysis [here](https://www.kaggle.com/headsortails/tidy-titarnic) to have a better idea. However, bear in mind that it is **just an example** and you will not be required to have that level of complexity in your project.

# Task 2: Exploring your dataset (15 points)

If we rewind and go back to the learning objectives, you'll see that by the end of this deliverable, you should have formulated *4* research questions about your data that you may want to answer during your project. However, it may be handy to do some more exploration on your dataset of choice before creating these questions - by looking at the data, you may get more ideas. **Before you start this task, read all instructions carefully until you reach START HERE under Task 3**.

2.1 Complete *4 out of the following 8 exercises* to dive deeper into your data. All datasets are different and therefore, not all of these tasks may make sense for your data - which is why you should only answer *4*. Use *dplyr* and *ggplot*.

1.  Plot the distribution of a numeric variable.
2.  Create a new variable based on other variables in your data (only if it makes sense)
3.  Investigate how many missing values there are per variable. Can you find a way to plot this?
4.  Explore the relationship between 2 variables in a plot.
5.  Filter observations in your data according to your own criteria. Think of what you'd like to explore - again, if this was the `titanic` dataset, I may want to narrow my search down to passengers born in a particular year...
6.  Use a boxplot to look at the frequency of different observations within a single variable. You can do this for more than one variable if you wish!
7.  Make a new tibble with a subset of your data, with variables and observations that you are interested in exploring.
8.  Use a density plot to explore any of your variables (that are suitable for this type of plot).

2.2 For each of the 4 exercises that you complete, provide a *brief explanation* of why you chose that exercise in relation to your data (in other words, why does it make sense to do that?), and sufficient comments for a reader to understand your reasoning and code.

<!-------------------------- Start your work below ---------------------------->
Data set:*apt_buildings*

**4 Exercise to explore the data:**

**1.  Plot the distribution of a numeric variable.**
```{r}
apt_buildings %>% ggplot(aes(x=year_built)) + geom_histogram(binwidth=.5, colour="black", fill="white")
```

-   *Brief explanation:* "year_built" is a numeric variable in this data set thus it is possible to plot a distribution of this variable. In regarding to the question, it would be interested to know when most of the buildings was built. IN this case, it seems that most of the building in vancouver was built in the 60s.

**2.  Explore the relationship between 2 variables in a plot.**
```{r}
apt_buildings %>% ggplot(aes(x=year_built,fill=air_conditioning)) + geom_density(alpha=.3)
```

-   *Brief explanation:* I am interested in knowing whether the more recent built building are more likely to have air-conditioning. However, there are multiple types of air-conditioning, so this type of density plot, would be the best to show us is the type of air conditioning (if any) that is used overtime.

**3.  Filter observations in your data according to your own criteria.**
```{r}
apt_buildings_after1900 <- apt_buildings %>% filter(year_built>1900)
dim(apt_buildings_after1900)
```

-   *Brief explanation:* Since buildings that are over 100 year old should be torn down soon or underwent major renovation that changes the structure and amenities (including the type of air conditionning), I will filter the building built after 1900.

**4.  Use a boxplot to look at the frequency of different observations within a single variable.**
```{r}
apt_buildings %>% ggplot(aes(x=air_conditioning,y=no_of_units)) + geom_boxplot()
apt_buildings %>% ggplot(aes(x=air_conditioning,y=no_of_units)) + geom_boxplot() + ylim(0,1000)
```

-   *Brief explanation:* I want to know if there is a relationship between number of units and the type of air_conditioning that a building have. Intuitively, the larger building should more likely to have central air conditioning. After trying the first line of code, there is an outlier with over 4000 units with the "Central air conditioning" so I rescale it in order to have a better look at other observations. After resealing, we can see that overall it seems that there might be a correlation between the type of air conditioning and number of units.

<!----------------------------------------------------------------------------->

# Task 3: Write your research questions (5 points)

So far, you have chosen a dataset and gotten familiar with it through exploring the data. Now it's time to figure out 4 research questions that you would like to answer with your data! Write the 4 questions and any additional comments at the end of this deliverable. These questions are not necessarily set in stone - TAs will review them and give you feedback; therefore, you may choose to pursue them as they are for the rest of the project, or make modifications!

<!--- *****START HERE***** --->
Data set:*apt_buildings*

-   Research question 1: Are newer buildings more likely to have barrier-free accessibility entry?

-   Research question 2: Is there a relationship between the type of visitor parking available and the size of building in term of number of units?

-   Research question 3: Is a building that allows smoking more or less likely to have a sprinkler system?

-   Research question 4: Is there any relationship between the type of window and the type of air conditioning system that is used?
<!----------------------------------------------------------------------------->

# Task 4: Process and summarize your data (13 points)

From Task 2, you should have an idea of the basic structure of your dataset (e.g. number of rows and columns, class types, etc.). Here, we will start investigating your data more in-depth using various data manipulation functions.

### 1.1 (10 points)

Now, for each of your four research questions, choose one task from options 1-4 (summarizing), and one other task from 4-8 (graphing). You should have 2 tasks done for each research question (8 total). Make sure it makes sense to do them! (e.g. don't use a numerical variables for a task that needs a categorical variable.). Comment on why each task helps (or doesn't!) answer the corresponding research question.

Ensure that the output of each operation is printed!

**Summarizing:**

1.  Compute the *range*, *mean*, and *two other summary statistics* of **one numerical variable** across the groups of **one categorical variable** from your data. (Done)
2.  Compute the number of observations for at least one of your categorical variables. Do not use the function `table()`! (Done)
3.  Create a categorical variable with 3 or more groups from an existing numerical variable. You can use this new variable in the other tasks! *An example: age in years into "child, teen, adult, senior".* (Done)
4.  Based on two categorical variables, calculate two summary statistics of your choosing. (Done)

**Graphing:**

5.  Create a graph out of summarized variables that has at least two geom layers. (Done)
6.  Create a graph of your choosing, make one of the axes logarithmic, and format the axes labels so that they are "pretty" or easier to read.(Done)
7.  Make a graph where it makes sense to customize the alpha transparency.(Done)
8.  Create 3 histograms out of summarized variables, with each histogram having different sized bins. Pick the "best" one and explain why it is the best.

Make sure it's clear what research question you are doing each operation for!

<!------------------------- Start your work below ----------------------------->
*Research question 1:* Are newer buildings more likely to have barrier-free accessibility entry?
**Summarizing:**  Create a categorical variable with 3 or more groups from an existing numerical variable. You can use this new variable in the other tasks!
```{r}
apt_buildings %>% select(year_built,barrier_free_accessibilty_entr) %>% mutate(year_built_r_cat=case_when(
      year_built< 1900 ~ "Before 1900",
       year_built> 2000 ~ "After 2000",
      TRUE                   ~ "Between 1900 and 2000"))
```

-   Comment: This task allows me to categorize the buildings' built years into 3 categorical variables from the existing numerical year built variable. However, this task alone will not help me answer the research question without adding additional line of code. However, it sets up so that I can compare the proportion of building with barrier-free accessibility entry between these 3 periods if that is something I want to explore.
**Graphing:**  Make a graph where it makes sense to customize the alpha transparency.
```{r}
apt_buildings %>% ggplot(aes(x=year_built, fill=barrier_free_accessibilty_entr)) +geom_density(alpha =0.5) +xlab('Year built')+ theme_minimal()
```

-   Comment: Since there are overlapping between group of variables over the year, it makes sense to customize alpha transparency. Examining this graph, it seems clear to me that newer building are more likely to have barrier-free entry compared to older building. In fact, most building built after the year 2000 have barrier-free accessibility entry.

*Research question 2:* Is there a relationship between the type of visitor parking available and the size of building in term of number of units?
**Summarizing:**  Compute the *range*, *mean*, and *two other summary statistics* of **one numerical variable** across the groups of **one categorical variable** from your data.
```{r}
apt_buildings %>% group_by(visitor_parking) %>% summarise (mean=mean(no_of_units), n = n())
apt_buildings %>% group_by(visitor_parking) %>% summarise (range=range(no_of_units), n = n())
apt_buildings %>% group_by(visitor_parking) %>% summarise (min=min(no_of_units), n = n())
apt_buildings %>% group_by(visitor_parking) %>% summarise (max=max(no_of_units), n = n())
```

-   Comment: This task provides me with a few types of summary statistics for buildings with each type of visitor parking (categorical variable), including mean, min, max, and the range of the number of units (numerical variable) for building with each type of visitor parking. If we classify a building by their average number of units, then it seems that smaller building are more likely to not have visitor parking available as the mean number of units for those building are the lowest. Thus, this task somewhat allows me to answer the research question.

**Graphing:**  Create a graph out of summarized variables that has at least two geom layers.
```{r}
apt_buildings %>% ggplot(aes(x=visitor_parking,y=no_of_units))+stat_summary(fun.y="mean", geom="bar")+  xlab('Type of visitor parking')+  ylab('Average Number of units')
```


-   Comment: This graph provides me the same information as the summarizing task, as it graphs the average number of units by type of visitor parking. Thus, similar to the other task, this task also allows me to answer this research question.

*Research question 3:* Is a building that allows smoking more or less likely to have a sprinkler system?
**Summarizing:**  Compute the number of observations for at least one of your categorical variables. Do not use the function `table()`!
```{r}
apt_buildings %>% count(`non-smoking_building`)
apt_buildings %>% count(sprinkler_system)
apt_buildings %>% select(`non-smoking_building`,sprinkler_system) %>%filter(`non-smoking_building`=="YES",sprinkler_system=="YES")
apt_buildings %>% select(`non-smoking_building`,sprinkler_system) %>%filter(`non-smoking_building`=="NO",sprinkler_system=="NO")
```

-   Comment: Both whether a building is non-smoking and whether the building has a sprinkle system are categorical variables.This tasks tell me that there are 2071 building that allowed smoking, among them, at least 610 does not have a sprinkle system. Similarly, there are 1290 buildings that do not allow smoking, and among the, at least 867 have sprinkle systems. I will need to conduct additional analysis to answer the research question.

**Graphing:**  Create a graph out of summarized variables that has at least two geom layers.
```{r}
apt_buildings %>% ggplot(aes(x = `non-smoking_building`, fill=sprinkler_system)) +  geom_bar() + theme_minimal() + xlab("Non-Smoking Building?")
```

-   Comment: This graph shows me that there are more buildings that allow smoking than non-smoking. The proportion of building with and without a sprinkle system seems to be about the same for smoking and non-smoking building. Without doing hypothesis testing, it will be impossible for me to asnwer the research question

*Research question 4:* Is there any relationship between the type of window and the type of air conditioning system that is used?
**Summarizing:**  Compute the number of observations for at least one of your categorical variables. Do not use the function `table()`!
```{r}
apt_buildings %>% count(window_type)
apt_buildings %>% count(air_conditioning)
```

-   Comment: The result of this task showed me that double panel is the most common type of window, while most buildings have no air conditioning. however, this does not afford me enough information to answer the research question. 

**Graphing:**  Create a graph of your choosing, make one of the axes logarithmic, and format the axes labels so that they are "pretty" or easier to read.
```{r}
apt_buildings %>% ggplot(aes(x = air_conditioning, fill=window_type)) + scale_y_log10("Counts Per log 10 scale")+  geom_bar() + theme_minimal() + xlab("Type of air conditioing")
```

-   Comment: Type of window and air conditioning are both categorical variables. The number of units with no air conditions vastly outnumbers other type of air conditions. Thus using a log 10 scale here make it easier to see more proportional difference and compared the type of window type for each type of air conditioning. Overall it seems that buidling without air-conditioning is more likely to be equipped single panel window, but I cannot be sure without doing hypothesis testing.

<!----------------------------------------------------------------------------->

### 1.2 (3 points)

Based on the operations that you've completed, how much closer are you to answering your research questions? Think about what aspects of your research questions remain unclear. Can your research questions be refined, now that you've investigated your data a bit more? Which research questions are yielding interesting results?

<!-------------------------- Start your work below ---------------------------->
*Research question 1*: Are newer buildings more likely to have barrier-free accessibility entry?

-   Comment: This research question can be answered with high degree certainty based on completed tasks, newer buildings are more likely to have to have 
barrier-free accessibility entry.

*Research question 2*: Is there a relationship between the type of visitor parking available and the size of building in term of number of units?

-   Comment: This research question can be answered with moderate degree certainty based on completed tasks, smaller building by number of units is less likely to have visitor parking.

*Research question 3*: Is a building that allows smoking more or less likely to have a sprinkler system?

-   Comment: We cannot currently answer this questions with certainty based on the based on completed tasks. We need to conduct hypotheis testing.

*Research question 4*: Is there any relationship between the type of window and the type of air conditioning system that is used?

-   Comment: We cannot currently answer this questions with certainty based on the based on completed tasks. We need to conduct hypotheis testing.

<!----------------------------------------------------------------------------->

### Attribution

Thanks to Icíar Fernández Boyano for mostly putting this together, and Vincenzo Coia for launching.
